---
title: "SF Forecasting"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE}
# Database Connect --------------------------------------------------------

## Install Packages

### Redshift

if(!require(DBI)) {
  install.packages("DBI")
  library(RPostgres)
}

if(!require(RPostgreSQL)) {
  install.packages("RPostgreSQL")
  library(RPostgreSQL)
}

if(!require(redshiftTools)) {
  install.packages("redshiftTools")
library(redshiftTools)
}
  
library("aws.s3")

### Forecast

if(!require(lubridate)) {
  install.packages("lubridate")
  library(lubridate)
}

if(!require(forecast)) {
  install.packages("forecast")
  library(forecast)
}

if(!require(xts)){
  install.packages("xts")
  library(xts)
}

if(!require(tseries)){
  install.packages("tseries")
  library(tseries)
}

### General

if(!require(tidyverse)){
  install.packages("tidyverse")
  library(tidyverse)
}

# Functions ---------------------------------------------------------------

create_test_train <- function(ts, h = 1){
  dates = as.Date(time(ts))
  min_date = min(dates)
  max_date = max(dates)
  list(holdout = tail(ts, h),
       train = ts(ts,
                  frequency = 12,
                  start = c(year(min_date),
                            month(min_date)),
                  end = c(year(max_date - months(h)),
                          month(max_date - months(h)))))
}


# Data Import -------------------------------------------------------------

# Load credentials for your own Redshift login
file <- paste("Z:\\User Output\\sf_odavies\\R WD\\R to Redshift\\my_credentials.csv")
my_credentials <- read.csv(file, header = TRUE, sep = ",")

# Load the credentials into variables.
port1 <- as.character(my_credentials[1,1])
host1 <- as.character(my_credentials[1,2])
user1 <- as.character(my_credentials[1,3])
password1 <- as.character(my_credentials[1,4])
dbname1 <- as.character(my_credentials[1,5])

# Establish a connection with Redshift
conn <- dbConnect(RPostgres::Postgres(), 
                        host=host1, 
                        port=port1, 
                        user=user1, 
                        password=password1, 
                        dbname=dbname1, 
                        sslmode='require')

## Type SQL Query - Sum of Sales per Month per Customer

query1 <- paste(
  "select
      left(cal_submit_date, 4) || '-' || substring(cal_submit_date, 6, 2) || '-28' as date,
      sum(fulfilled_sales) as sales
   from sf_analytics.prod_order
   where 
      cal_submit_date between dateadd(month, - 48, '2018-04-01') and '2018-03-31'
      and rp_person_id > 0
      and (kf_trade_split_validated = 'NTS'
      or kf_trade_split_validated is null)
   group by 1;"
)

```

# Introduction

This report details time series analysis of sales within NTS customer group.The main aim here is to create a forecast model that can extrapolate 15 months worth of sales. To achive this aim, we'll be using the Autoregressive Integrated Moving Average (ARIMA) model. 

The data are first sourced from the Redpoint:

```{r}
### Grab data
dat <- dbGetQuery(conn, query1)

### Convert string and order by date

dat$date <- ymd(dat$date)

dat <- dat %>%
        arrange(date)

### Max and Min Date

min_date <- min(dat$date)
max_date <- max(dat$date)

head(dat)
```

After we re-order the data by date and extract the minimum and maximum date for our forecasting analysis. We can then represent this series in R using  `ts()`:
```{r}
sf_nts_ts <- ts(data = dat$sales,
                start = c(year(min_date), month(min_date)),
                end = c(year(max_date), month(max_date)),
                frequency = 12)

sf_nts_ts
```

Notice the function requires start date (in `c(year, month)` format) and frequency which is always number of obvservations in a year.

To visualise the data we plot the time series:

```{r}
# Raw sales plot - notice trend/seasonality/variance
plot.xts(as.xts(sf_nts_ts/1E6),
         type = "o",
         ylab = "Sum of Sales [£m]",
         xlab = "Date",
         main = "NTS Sales Over Time")
```

We notice a positive trend with sales increasing over time. Seasonality (repeated periodic trends) is also apparent with characteristic dips across Dec-Feb and peaks across Oct-Nov. We can use time-series decompostion to examine these features using `stl()`  

```{r}
autoplot(stl(sf_nts_ts/1E6,
             s.window = 'periodic')) + geom_point()
```

These four plots represent the original data, the trend, the seasonality and the variance. These factors will be important to consider as we commonly assume in many time series techniques (especially ARIMA) that the time series is stationary. 

A stationary time series is one whose statistical properties such as mean, variance and correlation are all constant over time (i.e. to create the stationary time series, we need to minimise the trend, seasonality and variance). To achieve this we can apply differencing to our data or take logs.

